---
title: "National Park visits - US"
author: "Anna Reynolds"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  rmarkdown::html_document:
    code_folding: show
    theme: sandstone
    code_download: true
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(printr)
```

## Overview
This week's data is from dataisplural/data.world.

There are some additional datasets that may be interesting in relation to this data - namely US population and gas prices over time.

There's also some shapefiles you might be able to use with the sf package.

## Load data
```{r loading data, warning=FALSE, message=FALSE}
library(tidyverse)
library(sf)
library(ggrepel)
library(ggsn)

#Import data
park_visits <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/national_parks.csv")
state_pop <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/state_pop.csv")
gas_price <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/gas_price.csv")
head(park_visits)
#National Park centroids
park_cent <- read_sf(dsn = here::here("nps_boundary_centroids", "nps_boundary_centroids.shp"), layer = "nps_boundary_centroids")
#California counties
ca_count <- read_sf(dsn = here::here("CA_Counties", "CA_counties_TIGER2016.shp"), layer = "CA_counties_TIGER2016")
```

## Analysis/visualization

```{r map, }
park_ca <- park_cent %>% filter(STATE == "CA" & UNIT_TYPE == "National Park")
park_ca %>% ggplot() + geom_sf(data=ca_count) +
  geom_sf(data = park_ca, color = "dark green") + theme_minimal() +
  ggrepel::geom_label_repel(
    data = park_ca,
    aes(label = PARKNAME, geometry = geometry),
    stat = "sf_coordinates",
    min.segment.length = 0
  ) + labs(title = "National Parks, California", x = NULL, y = NULL) + north(park_ca, location = "topright", symbol = 3) + theme(axis.text = element_blank(),
                       axis.title = element_blank(),
                       axis.ticks = element_blank())
```

## Results


## Discussion/conclusions

## Package citations
  Hadley Wickham (2017). tidyverse: Easily Install and Load the
  'Tidyverse'. R package version 1.2.1.
  https://CRAN.R-project.org/package=tidyverse
  
  Pebesma, E., 2018. Simple Features for R: Standardized Support for
  Spatial Vector Data. The R Journal 10 (1), 439-446,
  https://doi.org/10.32614/RJ-2018-009
  
  Kamil Slowikowski (2019). ggrepel: Automatically Position Non-Overlapping
  Text Labels with 'ggplot2'. R package version 0.8.1.
  https://CRAN.R-project.org/package=ggrepel
  
  Oswaldo Santos Baquero (2019). ggsn: North Symbols and Scale Bars for
  Maps Created with 'ggplot2' or 'ggmap'. R package version 0.5.0.
  https://CRAN.R-project.org/package=ggsn
  
```{r citation, echo=FALSE, eval=FALSE}
citation("tidyverse")
citation("sf")
citation("ggrepel")
citation("ggsn")
```

